---
title: "Air pollution."
summary: "Health and climate change."
author: 'Giacomo Baldo'
date: '2021-10-24' 
slug: ''
categories: ["R"]
tags: ["Data Science", "Air pollution"]
image:
  caption: 'Credit: [https://climatekids.nasa.gov/air-pollution/](https://climatekids.nasa.gov/air-pollution/)'
  focal_point: ''
  preview_only: no
output:
  blogdown::html_page:
    df_print: kable  #kable, tibble, and paged are also options
---

# <mark> What is air pollution?</mark>

**When talking about air pollution, scientists refer to particles of solid, liquid and certain gases that are trapped in the air. ** 

**Air pollution has an impact on human health and climate, as well as visibility, biogeochemical cycling and atmospheric reactivity. **

```{r}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

```{r}
library(tidyverse)         # for graphing and data cleaning
library(kableExtra)        # for better tables
library(scales)            # to decide between scientific and extended number notation in plots
```

**Step 1)** Import the data. The data was previously downloaded to my laptop from the IEA website, the International Energy Agency.
```{r}
#read the dataset into “R”. The dataset will be called "data".
data <- read.csv("transport-sector-co2-emissions.csv")
``` 

**Step 2)** Display the data. 
```{r}
q = kbl(data) %>%
   kable_styling(bootstrap_options = c("responsive"))

q
```

**Step 3)** Display the data. 
```{r}
data_long <- data %>%
       pivot_longer(!Year, names_to = "Transport_type", values_to = "Emissions")

q1 <- kbl(data_long) %>%
            kable_styling(bootstrap_options = c("responsive"))

q1
``` 

```{r}

data_long$Transport_type <- factor(data_long$Transport_type, levels=c("Total","Passenger.road.vehicles", "Road.freight.vehicles", "Aviation", "Shipping", "Other","Rail" ), labels = c("Total","Psngr.road.veh.", "Road.freight.veh.", "Aviation", "Shipping", "Other","Rail" ) )

ggplot(data=data_long,  
        aes(x=Year, y= Emissions, colour=Transport_type)) +
        geom_line()+
   theme_light(base_size = 18) +
   theme(panel.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"), plot.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA")) +
   scale_y_continuous(name = "Greenhouse gases (Gigatons)", labels = comma) 
 ```
 

<!-- **Step 3)** Create and display a data set made of less information, recording only date, country and deaths. -->
<!-- ```{r} -->
<!-- data_less <- data[,c("date","deaths","countries")] -->

<!-- kbl(data_less[c(1:3,5003:5005),]) %>% -->
<!--    kable_styling(bootstrap_options = c("responsive")) -->
<!-- ``` -->

<!-- **Step 4)** Rearrange the data in a table with more columns using the function **pivot_wider()**. Then, display the first three rows.  -->
<!-- ```{r} -->
<!-- q <- data_less %>%  -->
<!--         pivot_wider(id_cols = date, -->
<!--                  names_from = countries, -->
<!--                 values_from = deaths) -->
<!-- kbl(q[c(1:3),]) %>% -->
<!--    kable_styling(bootstrap_options = c("responsive")) -->
<!-- ``` -->
<!-- With this new table, we can now readily compare the number of deaths across countries on a given date.  -->
