---
title: "Air pollution."
summary: "Health and climate change."
author: 'Giacomo Baldo'
date: '2021-10-24' 
slug: ''
categories: ["R"]
tags: ["Data Science", "Air pollution"]
image:
  caption: 'Credit: [https://climatekids.nasa.gov/air-pollution/](https://climatekids.nasa.gov/air-pollution/)'
  focal_point: ''
  preview_only: no
output:
  blogdown::html_page:
    df_print: kable  #kable, tibble, and paged are also options
---

# <mark> What is air pollution?</mark> #

**When talking about air pollution, scientists refer to particles of solid, liquid and certain gases that are trapped in the air. ** 

**Air pollution has an impact on human health and climate, as well as visibility, biogeochemical cycling and atmospheric reactivity. **

```{r}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

```{r}
library(tidyverse)         # for graphing and data cleaning
library(kableExtra)        # for better tables
library(scales)            # to decide between scientific and extended number notation in plots
library(ggrepel)
library(grid)
```

# <mark> Greenhouse gas emissions</mark> #

![](ghg-emissions-by-sector.svg)

**Step 1)** Import the data. The data was previously downloaded to my laptop from the IEA website, the International Energy Agency.
```{r}
#read the dataset into “R”. The dataset will be called "data".
data <- read.csv("transport-sector-co2-emissions.csv")
``` 

**Step 2)** Display the data. 
```{r}
q = kbl(data) %>%
   kable_styling(bootstrap_options = c("responsive"))

q
```

**Step 3)** Display the data. 
```{r}
data_long <- data %>%
       pivot_longer(!Year, names_to = "Transport_type", values_to = "Emissions")

q1 <- kbl(data_long) %>%
            kable_styling(bootstrap_options = c("responsive"))

head(q1)
``` 

```{r}

data_long$Transport_type <- factor(data_long$Transport_type, levels=c("Total","Passenger.road.vehicles", "Road.freight.vehicles", "Aviation", "Shipping", "Other","Rail" ), labels = c("Total","Passenger road \n vehicles", "Road freight \n vehicles", "Aviation", "Shipping", "Other","Rail" ) )

tp = ggplot(data=data_long,aes(x=Year, y= Emissions, colour=Transport_type)) +
      geom_line() +
  geom_text_repel(data = subset(data_long, Year == "2018"), aes(x=Year, y= Emissions, label = subset(data_long, Year == "2018")$Transport_type), segment.color = NA, size = 5) + ggtitle("Greenhouse gas emissions by transport sector")+
  scale_colour_discrete(guide = 'none')  +    
  theme_light(base_size = 15) +   
  theme(plot.margin = unit(c(1,1,1,1), "lines")) +
   theme(panel.background = element_rect(fill = "#FFFFFF", colour = "#FFFFFF"), plot.background = element_rect(fill = "#FFFFFF", colour = "#FFFFFF")) +
   scale_y_continuous(name = "Gigatons") 

tp
```      
   
         
