---
title: R for Data Science
summary: Data wrangling
author: 'Giacomo Baldo'
date: '2021-08-31' 
slug: ''
categories: ["R"]
tags: ["R", "Data Science", "Data wrangling", "pivot_wider()"]
image:
  caption: ''
  focal_point: ''
  preview_only: no
output:
  blogdown::html_page:
    df_print: kable  #kable, tibble, and paged are also options
---
# <mark>Data wrangling</mark> with <mark>pivot_wider()</mark>.

**Step 0)** This page starts by setting up the options for this document and loading the R libraries required later.
```{r}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

```{r}
library(tidyverse)         # for graphing and data cleaning
library(kableExtra)        # for better tables
library(scales)            # to decide between scientific and extended number notation in plots
```

**Step 1)** Import the data. The data was previously downloaded to my laptop (on 18/08/2021) from the ECDC website, the European Centre for Disease prevention and Control website.
```{r}
#read the dataset into “R”. The dataset will be called "data".
data <- read.csv("covid-data.csv")
```

**Step 2)** Display six arbitrarily chosen rows of the data, in this case rows 1 to 3 and 5003 to 5005.
```{r}
kbl(data[c(1:3,5003:5005),]) %>%
   kable_styling(bootstrap_options = c("responsive"))
```

**Step 3)** Create and display a data set made of less information, recording only date, country and deaths.
```{r}
data_less <- data[,c("date","deaths","countries")]

kbl(data_less[c(1:3,5003:5005),]) %>%
   kable_styling(bootstrap_options = c("responsive"))
```

**Step 4)** Rearrange the data in a table with more columns using the function **pivot_wider()**. Then, display the first three rows. 
```{r}
q <- data_less %>% 
        pivot_wider(id_cols = date,
                 names_from = countries,
                values_from = deaths)
kbl(q[c(1:3),]) %>%
   kable_styling(bootstrap_options = c("responsive"))
```
With this new table, we can now readily compare the number of deaths across countries on a given date. 




#### I'll be refining and adding to this analysis in the near future.
 
