---
title: R for Data Science
summary: Using R to explore Covid-19
author: 'Giacomo Baldo'
date: '2021-08-17'
slug: ''
categories: ["R"]
tags: ["R", "Data Science"]
image:
  caption: ''
  focal_point: ''
  preview_only: no
output:
  blogdown::html_page:
    df_print: kable  #kable, tibble, and paged are also options
---
 
```{r include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning=FALSE)
```

```{r, include=FALSE,  warning=FALSE}
#library(lubridate)         # for working with dates
#library(palmerpenguins)    # for palmer penguin data
#library(remotes)         # for installing package from GitHub
#remotes::install_github("llendway/gardenR") # run if package is not already installed
#library(gardenR)           # for Lisa's garden data
#theme_set(theme_minimal()) # my favorite ggplot theme
```

# Using <mark>ggplot()</mark> and <mark>dplyr</mark >

In this section I take a look at data related to the Covid-19 epidemic. The data is imported from the European Centre for Disease Prevention and Control (ECDC) website. First, I show a table for the cumulative deaths and cases for each EU country for which data is available. Numbers are usually up to date.

Before reading on, ask yourself "which EU country was the hardest hit by Covid-19?" and "is the wealth of countries a good predictor of deaths due to Covid-19?"

&nbsp;

## Cumulative deaths and cases for each country.

We look at the total number of deaths for each country for which data is available from the ECDC. First, the data is displayed in a table. Below the table, the same data is displayed in a bar plot.

```{r}
library(tidyverse)         # for graphing and data cleaning
library(kableExtra)        # for better tables
library(scales)

#read the dataset into “R”. The dataset will be called "data".
data <- read.csv("covid-data.csv")

#extract cumulative deaths and cases for each EU country for which data is available
q <- data %>% 
  group_by(countries) %>% 
  summarize(deaths = sum(deaths),
            cases = sum(cases)) %>% 
  arrange(desc(deaths))

#plot the table
kbl(q) %>%
   kable_styling(bootstrap_options = c("responsive"))

q$countries <- factor(q$countries, levels = q$countries[order(q$deaths)])

z <- ggplot(data = q, mapping = aes(x = deaths, y = countries)) +
  geom_bar(stat='identity') +
  theme_light() +
  theme(panel.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"), plot.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA")) +
  scale_x_continuous(name = "Cumulative deaths", labels = comma)

z
```
&nbsp; 

## Deaths per million inhabitants

Although Italy was the hardest hit by the virus in absolute values, it was not in terms of deaths per million. The next plot shows this. 
```{r}
library(tidyverse)
library(kableExtra)
library(scales)

data <- read.csv("covid-data.csv")

q <-  data %>% 
  group_by(countries,population) %>% 
  summarize(deaths = sum(deaths))

q <- tibble(countries = q$countries, deaths_pm = q$deaths/q$population*1000000) %>%
  arrange(desc(deaths_pm)) 

q$countries <- factor(q$countries, levels = q$countries[order(q$deaths_pm)])

#create a bar plot of the data above
z <- ggplot(data = q, mapping = aes(x = deaths_pm, y = countries)) +
  geom_bar(stat='identity') +
  theme_light() +
  theme(panel.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"), plot.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA")) +
  scale_x_continuous(name = "Deaths per million people", labels = comma)

#show the plot
z
```
&nbsp;

## Deaths per million people as a function of per capita GDP

Finally, a trend seems to emerge in which countries with a low per-capita GDP tend to having more deaths per million people.
```{r}
library(tidyverse)
library(kableExtra)
library(scales)
library(ggrepel)

data <- read.csv("covid-data.csv")
gdp_data <- read.csv("gdp-europe.csv")
merged_data <- merge(data, gdp_data, by = "countries")

q <- merged_data %>% 
  group_by(countries,gdp_pp,population) %>% 
  summarize(deaths = sum(deaths))

q <- tibble(countries = q$countries, gdp_pp = q$gdp_pp, deaths_pm = q$deaths/q$population*1000000)

t = ggplot(data = q, mapping = aes(x = gdp_pp, y = deaths_pm)) +
  geom_point() + 
  geom_text_repel(aes(label=countries)) +
  theme_light() +
  scale_x_continuous(name = "GDP per capita (€)") +
  scale_y_continuous(name = "Deaths per million people", labels = comma) +
  theme(panel.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"), plot.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"))

t 
```
&nbsp;

## Linear regression 

Now it seems the right time to check for a linear relationship between deaths and GDP. A preliminary analysis has shown that removing the countries with a population smaller than one million from the data set improves the p-values of the regression method.

In the next bit of analysis, first I identify and remove these countries from the data set. Then, I perform the linear regression on the new data set. 

```{r}
library(tidyverse)
library(kableExtra)
library(scales)
library(ggrepel)

data <- read.csv("covid-data.csv")
gdp_data <- read.csv("gdp-europe.csv")
merged_data <- merge(data, gdp_data, by = "countries")

#countries with a population size smaller than one million
c_one_m <- unique(merged_data[which(merged_data$population < 1000000), ]$countries)
c_one_m <- paste(c_one_m, collapse = ", ")
print(paste(" The countries with a population size less than one million are", c_one_m, sep = " "))

#new data set with only countries with a population size over one million
merged_data <- merged_data[(merged_data$population > 1000000), ]

#create a table 
q <- merged_data %>% 
  group_by(countries,gdp_pp,population) %>% 
  summarize(deaths = sum(deaths))

#new table with added column "deaths per million people"
q <- tibble(countries = q$countries, gdp_pp = q$gdp_pp, deaths_pm = q$deaths/q$population*1000000)

#create plot
t <- ggplot(data = q, mapping = aes(x = gdp_pp, y = deaths_pm)) +
  geom_point() + 
  geom_text_repel(aes(label=countries)) +
  theme_light() +
  scale_x_continuous(name = "GDP per capita (€)") +
  scale_y_continuous(name = "Deaths per million people", labels = comma) +
  theme(panel.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"), plot.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"))


#add to the graph the line of best fit derived from the linear regression
u <- t +
  geom_smooth(method='lm', formula= y~x, se = FALSE)

u

#perform linear regression without plotting
m <- lm(deaths_pm ~ gdp_pp, # How does deaths_pm depend on (~) gdp_pp
        data = q) # Look in "q" for the variables

#print summary of linear regression
print(summary(m))

#print the q-q plot
ggplot() +
  stat_qq(aes(sample = m$residuals)) +
  stat_qq_line(aes(sample = m$residuals)) + 
  theme_light() +
  scale_x_continuous(name = "theoretical quantiles") +
  scale_y_continuous(name = "sample quantiles") +
  theme(panel.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA"),
plot.background = element_rect(fill = "#F8F9FA", colour = "#F8F9FA")) +
  ggtitle("Q-q plot") 

```
&nbsp;

## Comments on the line of best fit 

#### The line of best fit 

The "coefficients" columns in the summary output tell us that the line of best fit has the following equation:

$$  y = 2.414 \cdot 10^{-3} - 3.073 \cdot 10^{-2} \cdot x $$
which is equivalent to

$$  y = 0.002414 - 0.03073 \cdot x. $$

#### Interpreting the line of best fit 

The variable $x$ stands for GDP per capita, while $y$ stands for deaths per million people. The $x$ coefficient tells us that for an increase in €10,000 of per-capita GDP, deaths are expected to reduce by 307 (per million people).

## The supporting statistical evidence 

#### The p-value

The evidence for accepting this relationship is given by the very small p-value given for "gdp_pp" resulted in the linear regression summary statistics. The p-value is equal to 0.000265. In more technical terms (read on if interested in technicalities), this means that we can confidently reject the null-hypothesis (assuming the null-hypothesis to be true, there is only a 0.0265% probability of obtaining a test statistic equal or more extreme than the t-value). 

#### The distribution of residuals

The p-value of the standard linear regression performed above is meaningful if the residuals follow a normal distribution. In the summary statistics, one can see that residuals are symmetrically distributed with mean close to 0 (in relative terms). This is already good evidence for a normal distribution. A more rigorous test for normal distribution is given by a q-q plot, which compares what would be the residuals of data that follow a standard normal distribution to the residuals from the tested data. When the resulting q-q plot follows a straight line, as it roughly does in our case above, the distribution of the residuals does indeed closely follow a normal distribution.





## The end

This analysis points to the outcomes of economic inequalities. It shows a strong relationship between the wealth of a country and the ability to protect its citizens from Covid-19. 

Finally, thank you for visiting this page. I'll be refining and adding to this analysis in the near future.
 
